---
title: "Reproducible Research, assignment week 2"
author: "P.A. de Roos"
date: "February 7, 2017"
output: html_document
---


Set the working directory & load the dataset.

```{r loadTheData}
setwd("C:/Users/paul/OneDrive/Coursera/Reproducible Research")
Activity <- read.csv("activity.csv")
head(Activity)
```

Calculate the number of steps per day and display as histogram.

```{r CalculateSteps}
(StepsPerDay <- with(Activity, tapply(steps, date, sum)))
hist(StepsPerDay)
```

Calculate the mean and median of the total number of steps per day.

```{r CalculateMeanAndMedianOfSteps}
(MeanStepsPerDay <- mean(StepsPerDay, na.rm = T))
(MedianStepsPerDay <- median(StepsPerDay, na.rm = T))
```

Create a time series plot of average steps per time of day.

```{r TimeSeriesPlot}
StepsPerInterval <- with(Activity, tapply(steps, interval, mean, na.rm = T))
Intervals <- sort(unique(Activity$interval), decreasing = F)
StepsWithIntervals <- as.data.frame(cbind(Intervals, StepsPerInterval))
plot.ts(x=StepsWithIntervals$Intervals, y= StepsWithIntervals$StepsPerInterval, type="l")
```

Calculate the interval which has the highest amount of steps.

```{r HighestSteps}
(HighestSteps <- StepsWithIntervals[which(StepsWithIntervals$StepsPerInterval == max(StepsWithIntervals$StepsPerInterval)),])
```

Count the number of days with NA values (note these dates may also have non-NA values).

```{r CountNAs}
length(unique(Activity$date[is.na(Activity$steps)]))
```

Fill the NA values with the average steps per interval, step 1: create a new dataset without NA's.

```{r NoNAs}
Activity2 <- Activity[!is.na(Activity),]
```

Fill the NA values with the average steps per interval, step 2: calculate the average number of steps per interval.
```{r MeanStepsPerInterval}
MeanStepsPerInterval <- aggregate(Activity2$steps, list(Activity2$interval), mean, na.action = na.omit)
(colnames(MeanStepsPerInterval) <- c("interval", "meanSteps"))
```

Fill the NA values with the average steps per interval, step 3: merge the avg. no. of steps per interval with a copy of the original dataset.
```{r Merge}
library(data.table)
Activity3 <- Activity
Activity3 <- merge(Activity, MeanStepsPerInterval)
head(Activity3)
```

Fill the NA values with the average steps per interval, step 4: replace the NA's with the averages, then lose the meanSteps column.
```{r Replace}
Activity3$steps <- ifelse(!is.na(Activity3$steps) , Activity3$steps, Activity3$meanSteps)
Activity3 <- Activity3[-4]
head(Activity3)
```

Calculate the number of steps per day -based on the dataset with imputed values- and display as histogram.

```{r CalculateSteps_withImputedValues}
(StepsPerDay_withImputedValues <- with(Activity3, tapply(steps, date, sum)))
hist(StepsPerDay_withImputedValues)
```

...note there is hardly a difference in the chart compared to the previous histogram.

Calculate the mean and median of the total number of steps per day -based on the dataset with imputed values-.

```{r CalculateMeanAndMedianOfSteps_withImputedValues}
(MeanStepsPerDay_withImputedValues <- mean(StepsPerDay_withImputedValues, na.rm = T))
(MedianStepsPerDay_withImputedValues <- median(StepsPerDay_withImputedValues, na.rm = T))
```

Create a new factor variable (weekday/weekend).

```{r weekdays}
weekdayOrWeekend <- as.factor(weekdays(as.Date(Activity3$date)))
weekdayOrWeekend <- ifelse(weekdayOrWeekend=="Saturday"|weekdayOrWeekend=="Sunday", "Weekend", "Weekday")
Activity_WoW <- cbind(Activity3, weekdayOrWeekend)
head(Activity_WoW)
```

Create a time series plot of average steps per time of day, split per weekday/weekend.

Step 1: calculate the average steps per interval and attach that as an extra column to the Activity_WoW dataset.
NB WoW stands for "Weekday or Weekend".

```{r TimeSeriesPlot_weekdayOrWeekend_step1}
StepsPerInterval_WoW <- with(Activity_WoW, tapply(steps, interval, mean, na.rm = T))
Intervals_WoW <- sort(unique(Activity_WoW$interval), decreasing = F)
StepsWithIntervals_WoW <- as.data.frame(cbind(Intervals_WoW, StepsPerInterval_WoW))
```

Step 2: Split the dataset in 2 dataset and plot them both.

```{r TimeSeriesPlot_weekdayOrWeekend_step2}
Weekday <- StepsWithIntervals_WoW[weekdayOrWeekend =="Weekday",]
Weekend <- StepsWithIntervals_WoW[weekdayOrWeekend =="Weekend",]
par(mfrow=c(2,1))
plot.ts(x=Weekday$Intervals, y= Weekday$StepsPerInterval, type="l", main="Weekdays", xlab="Intervals", ylab="Steps per interval")
plot.ts(x=Weekend$Intervals, y= Weekend$StepsPerInterval, type="l", main="Weekend", xlab="Intervals", ylab="Steps per interval")

```






